Integrated nested Laplace approximations for extended latent Gaussian models with application to the Naomi HIV model

### 2023-01-18 src/docs_paper/paper.Rmd
Some Jeff and Seth comments

Mounting an effective public health response to the HIV epidemic requires accurate, timely HIV indicator estimates at a sufficiently fine-scale resolution to make targetted interventions.
Producing these estimates is a challenging task beccause all available data sources have shortcomings which must be overcome.
Nationally-representative household surveys provide the most statistically reliable data, but due to their high cost to run, in most countries they occur only infrequently.
Other data sources, such as routine health surveillance of antenatal care clinics, are more real-time but based on limited or biased samples of the population.
To meet these challenges, the Naomi small-area estimation model [@eaton2021naomi] synthesises data from multiple sources to estimate HIV prevalence, HIV incidence, and coverage of antiretroviral treatment (ART) at a district-level.
Simultaneous modelling of the data mitigates the limitations of any single source, increases statistical power, and prompts investigation into any conflicting information.
Software (\url{https://naomi.unaids.org}) has been developed for Naomi, allowing countries to input their data and interactively generate estimates in a yearly process supported by UNAIDS.
The creation of estimates by country teams, rather than external agencies, is a noteworthy feature of the HIV response.
Drawing on expertise closest to the data generating process, improves the accuracy of the process, as well as strengthening trust and ownership of the estimates.

The use case requirements for the model, combined with its relative complexity, present a difficult Bayesian inference problem.
Any inferential strategy must be fast, as well as easy to run in production by country teams, ruling out prohibitively slow Markov chain Monte Carlo (MCMC) approaches.
Inference is currently conducted using an empirical Bayes approach, with a Gaussian approximation to the latent field, via the Template Model Builder (`TMB`) \textsf{R} package [@kristensen2016tmb].
Owing to its speed and flexibility, `TMB` has recently been gaining popularity more broadly in spatial statistics [@osgoodzimmerman2021statistical].
Inference in `TMB` is based on optimisation of a \textsf{C++} template function, with the option available to use a Laplace approximation to integrate out any subset of the function arguments, which for the Naomi model, we use for the high-dimensional latent field parameters.
Taking inspiration from the AD Model Builder (ADMB) package [@fournier2012ad], `TMB` uses automatic differentiation [@baydin2017automatic] to calculate the derivatives required for numerical optimisation routines and the Laplace approximation.
Although this approach has favourable computational properties, we have found the inferences generated for the Naomi model to sometimes be inaccurate.
This has motivated us to look for a more accurate approach, which is flexible enough to be compatible with the model and fast enough to be run in production by country teams.

To obtain fast, accurate inferences for the Naomi model we develop a new inference methodology which combines the simplified integrated nested Laplace approximation (INLA) approach of @wood2020simplified with adaptive Gauss-Hermite quadrature (AGHQ).
INLA is an approach to approximate Bayesian inference based on nested Laplace approximations and numerical quadrature.
The central innovation of @rue2009approximate is an approximation which enables accurate latent field posterior marginals without explicitly computing the full Laplace approximation for each element.
Simplified INLA [@wood2020simplified] extends INLA by relaxing the sparsity assumptions on the latent field required for this approximation to be accurate.
This extension facilitates inference for models like Naomi, which fall within the extended latent Gaussian model (ELGM) [@stringer2022fast] class, and were not previously amenable to inference with INLA.
ELGMs build on latent Gaussian models (LGMs) by allowing each element of the linear predictor to depend on any subset of elements from the latent field.
We combine simplified INLA with AGHQ, a quadrature rule based on the theory of polynomial interpolation which adapts to the integrand based on the Hessian at the mode.
Though no theory yet exists for the nested case, the first stochastic convergence results for adaptive quadrature rules were recently obtained by @bilodeau2021stochastic using AGHQ.
We implement our method as an extension of the `aghq` \textsf{R} package [@stringer2021implementing].
Since `aghq` is designed to naturally interface with `TMB`, use of our method is simple when provided a \textsf{C++} user template for the log-posterior.

The remainder of this paper is organised as follows.
In Section \ref{sec:naomi} we describe a simplified version of the Naomi model that we consider in this paper.
Section \ref{sec:fastinferencemethods} outlines our approach to fast, accurate Bayesian inference using simplified INLA and AGHQ.
As a case-study, we fit the simplified Naomi model on data from Malawi, and compare the accuracy of inferences in Section \ref{sec:naomi}.
We also demonstrate a Bayesian workflow, illustrating the applicability of these tools in a deterministic inference setting.
Finally, in Section \ref{sec:conclusions} we discuss our conclusions, how we anticipate our method might be useful for other models, and directions for future research.

@@

@eaton2021naomi specify a joint model linking small-area estimation models of HIV prevalence from household surveys, HIV prevalence from antenatal care clinics, and antiretroviral therapy (ART) coverage from routine health data collection.
The model is defined over three time points: $T_1$ the time of the most recent household survey with HIV testing; $T_2$, the current time period; and $T_3$, a short term projection period.
We consider a simplified version of the model which is defined only at $T_1$ omitting temporal projection to $T_2$ and $T_3$.
Below we provide an overview of the three components in the simplified model, highlighting the aspects which make it a challenge for existing inferential approaches.
A more complete mathematical description of the simplified model, as well as a \textsf{C++} template for the log-posterior, are provided in the appendix.

@@

Consider a country in sub-Saharan Africa where a household survey has taken place during the quarter $T_1$.
Let $x$ index district, $a$ five-year age band, $s$ sex and $t$ time.
For ease of notation, let $i$ index the coarsest district-age-sex division included in the model.
The data we observe may be aggregated over indices $i$, so we let $\mathcal{I}$ be a set of $i$ for which observations are reported.
Let $N_i \in \mathbb{N}$ be the population size, $\rho_i \in [0, 1]$ be HIV prevalence, $\alpha_i \in [0, 1]$ be ART coverage, and $\lambda_i > 0$ be the annual HIV incidence rate.
We specify independent mixed effects models for HIV prevalence and ART coverage in the general population on the logit scale such that
\begin{align*}
\text{logit}(\rho_i) &= \eta^\rho_i, \\
\text{logit}(\alpha_i) &= \eta^\alpha_i,
\end{align*}
for certain choice of linear predictors $\eta^\rho_i$ and $\eta^\alpha_i$.
For the HIV incidence rate we use a mixed effects model on the log scale
$$
\log(\lambda_i) = \eta^\lambda_i(\{\rho_i, \alpha_i\}_{i \in \mathcal{I}}),
$$
where the linear predictor depends on $\{\rho_i, \alpha_i\}_{i \in \mathcal{I}}$ for some $\mathcal{I}$.
Let $\kappa_i$ be the proportion recently infected among HIV positive persons.
For each set of observed strata indices $\mathcal{I}$, we calculate the weighted observations $\hat \theta_\mathcal{I}$ for $\theta \in \{\rho, \alpha, \kappa\}$ with respective Kish effective sample sizes 
$$
M^{\hat \theta}_\mathcal{I} = \frac{\left(\sum_j w_j\right)^2}{\sum_j w_j^2},
$$
where $j$ indexes individuals across all strata $i \in \mathcal{I}$, with corresponding survey weights $w_j$.
The observed number of indicator cases is then
$$
Y^{\hat \theta}_{\mathcal{I}} = M^{\hat \theta}_{\mathcal{I}} \cdot \hat \theta_{\mathcal{I}}.
$$
For $\theta \in \{\rho, \alpha, \kappa\}$ we model these aggregate observations using a binomial working likelihood
$$
Y^{\hat \theta}_{\mathcal{I}} \sim \text{xBin}(M^{\hat \theta}_{\mathcal{I}}, \theta_{\mathcal{I}}),
$$
where $\theta_{\mathcal{I}}$ are the following appropriately weighted aggregates
\begin{align*}
\rho_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i}{\sum_{i \in \mathcal{I}} N_i}, \\
\alpha_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \alpha_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}, \\
\kappa_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \kappa_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}.
\end{align*}
To link the proportion recently infected among HIV positive persons $\kappa_i$ to HIV incidence $\lambda_i$ we use
$$
\kappa_i = 1- \exp \left( - \lambda \cdot \frac{1 - \rho_i}{\rho_i} \cdot (\Omega_T - \beta_T) - \beta_T \right),
$$
where the mean duration of recent infection $\Omega_T$ and  is the false recent ratio $\beta_T$ are strongly informed by priors for the particular survey.

@@

We model HIV prevalence $\rho^\text{ANC}_i$ and ART coverage $\alpha^\text{ANC}_i$ among pregnant women as being offset on the logit scale from the general population indicator as follows
\begin{align*}
\text{logit}(\rho^\text{ANC}_i) &= \text{logit}(\rho_i) + \eta^{\rho^\text{ANC}}_i, \\
\text{logit}(\alpha^\text{ANC}_i) &= \text{logit}(\alpha_i) + \eta^{\rho^\text{ANC}}_i.
\end{align*}
We inform these processes using the following aggregate ANC data from the year of the most recent survey: the number of ANC clients with ascertained status $X^\text{ANC}_\mathcal{I}$, the number of those with positive status $Y^\text{ANC}_\mathcal{I}$, and the number of ANC clients already on ART prior to their first ANC visit $Z^\text{ANC}_\mathcal{I}$.
We use the binomial working likelihoods
\begin{align*}
Y^\text{ANC}_\mathcal{I} &\sim \text{Bin}(Y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}) \\
Z^\text{ANC}_\mathcal{I} &\sim \text{Bin}(Y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}),
\end{align*}
where again we use weighted aggregates
\begin{align*}
\rho^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}}{\sum_{i \in \mathcal{I}} \Psi_i}, \\
\alpha^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho^\text{ANC} \alpha^\text{ANC}_i}{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}},
\end{align*}
with $\Psi_i$ the number of pregnant women.

@@

Given a \textsf{C++} user template `model.cpp` for the negative unnormalised log posterior $- \log p(\y, \x, \btheta)$, we obtain the posterior marginal approximations $\{ \tilde p(x_i \, | \, \y) \}_{i = 1}^n$ and $\tilde p(\theta_j \, | \, \y)_{j = 1}^m$ via the following algorithm, comprised of nested applications of Laplace approximation and adaptive Gauss-Hermite quadrature.

@@

1. Calculate
\begin{align*}
\hat{\btheta} &= \argmax_{\btheta} {\tilde p_\text{LA}(\btheta, \y)}, \\
\mathbf{H} &= \frac{\partial^2}{\partial \btheta \partial \btheta^\top} - \log \tilde p_\text{LA}(\btheta, \y) \rvert_{\btheta = \hat \btheta} \\
\mathbf{H}^{-1} &= \mathbf{L} \mathbf{L}^\top,
\end{align*}
where
$$
\tilde p_\text{LA}(\btheta, \y) = \frac{p(\y, \x, \btheta)}{\tilde p_\text{G}(\x \, | \, \btheta, \y)} \Big\rvert_{\x = \hat \x(\btheta)}
$$
where $\tilde p_\text{G}(\x \, | \, \btheta, \y) = \mathcal{N}(\x \, | \, \hat \x(\btheta), \mathbf{H}(\btheta)^{-1})$ is a Gaussian approximation to $p(\x \, | \, \btheta, \y)$ with mode and precision matrix given by
\begin{align*}
\hat \x(\btheta) &= \argmin_\x - \log p(\y, \x, \btheta), \\
\mathbf{H}(\btheta) &= \frac{\partial^2}{\partial x \partial x^\top} - \log p(\y, \x, \btheta) \rvert_{\x = \hat \x(\btheta)}.
\end{align*}
2. Generate a set of nodes $\z \in \mathcal{Q}(m, k)$ and weights $\omega: \z \in \mathcal{Q}(m, k) \to \mathbb{R}$ from a Gauss-Hermite quadrature rule with $m = \dim(\btheta)$, $k$ nodes per dimension, as follows
\begin{align*}
\btheta(\z) &= \hat{\btheta} + \mathbf{L} \z, \\
\hat{\btheta} &= \argmax_{\btheta} {\tilde p_\text{LA}(\btheta, \y)}, \\
\mathbf{H} &= \frac{\partial^2}{\partial \btheta \partial \btheta^\top} - \log \tilde p_\text{LA}(\btheta, \y) \rvert_{\btheta = \hat \btheta} \\
\mathbf{H}^{-1} &= \mathbf{L} \mathbf{L}^\top.
\end{align*}
We typically set $k = 3$ such that there are $3^m$ nodes in total.
3. Use this quadrature rule to calculate $\tilde p_{\text{AQ}}(\y)$ as follows
$$
\tilde p_{\text{AQ}}(\y) = \sum_{\z \in \mathcal{Q}(m, k)} \tilde p_\text{LA}(\btheta(\z), \y) \omega(\z).
$$
4. For $x_i \in \{\ldots\}$ and $\btheta \in \{ \btheta(\z) \}_{\z \in \mathcal{Q}(m, k)}$ calculate the modes and Hessians
\begin{align*}
\hat \x_{-i}(x_i, \btheta) &= \argmin_{\x_{-i}} - \log p(\y, x_i, \x_{-i}, \btheta), \\
\mathbf{H}_{-i, -i}(x_i, \btheta) &= \frac{\partial^2}{\partial \x_{-i} \partial \x_{-i}^\top} - \log p(\y, x_i, \x_{-i}, \btheta) \rvert_{\x_{-i} = \hat \x_{-i}(x_i, \btheta)},
\end{align*}
where optimisation to obtain $\hat \x_{-i}(x_i, \btheta)$ may be initialised at $\hat \x(\btheta)_{-i}$.

@@

* We developed an approximate Bayesian inference algorithm to solve a challenging problem in the small-area estimation of HIV in low resource settings
* Following further testing, we anticipate adding our inference method as an option for 
* The flexibility of our method implementation, including compatibility with any `TMB` \textsc{C++} template, allows broader use, as well as investigation of, deterministic inference methods than had previously been possible. We are excited about exploration of its accuracy for the following challenging model structures which could not previously be fit: list of model types here.
* We demonstrated a Bayesian workflow for deterministic inference methods

>>>>

Mounting an effective public health response to the HIV epidemic requires accurate, timely HIV indicator estimates at the level of geographic resolution at which heath systems are planned and delivered.
Producing these estimates is challenging because all available data sources have shortcomings which must be overcome.
Nationally-representative household surveys provide the most statistically reliable data, but due to their high cost, in most countries they occur only every five years or so, with limited sample size at the district level.
Other data sources, such as routine health surveillance of antenatal care clinics, are more real-time but based on limited or non-representative samples of the population.
To meet these challenges, the Naomi small-area estimation model [@eaton2021naomi] synthesises data from multiple sources to estimate HIV indicators at a district-level.
Simultaneous modelling of the data mitigates the limitations of any single source, increases statistical power, and prompts investigation into any conflicting information.
Software (\url{https://naomi.unaids.org}) has been developed for Naomi, allowing countries to input their data and interactively generate estimates in a yearly process supported by UNAIDS.
The creation of estimates by country teams, rather than external agencies, is a noteworthy feature of the HIV response.
Drawing on expertise closest to the data being modelled improves the accuracy of the process, as well as strengthening trust and ownership of the estimates.

The use case requirements for the model, combined with its relative complexity, present a difficult Bayesian inference problem.
Any inferential strategy must be fast enough for interactive review and iteration of modelling results, as well as easy to run in production by country teams, ruling out prohibitively slow Markov chain Monte Carlo (MCMC) approaches.
Inference is currently conducted using an empirical Bayes approach, with a Gaussian approximation to the latent field, via the Template Model Builder (`TMB`) \textsf{R} package [@kristensen2016tmb].
Owing to its speed and flexibility, `TMB` has recently been gaining popularity more broadly in spatial statistics [@osgoodzimmerman2021statistical].
Inference in `TMB` is based on optimisation of a \textsf{C++} template function, with the option available to use a Laplace approximation to integrate out any subset of the function arguments, which for the Naomi model, we use for the high-dimensional latent field parameters.
Taking inspiration from the AD Model Builder (ADMB) package [@fournier2012ad], `TMB` uses automatic differentiation [@baydin2017automatic] to calculate the derivatives required for numerical optimisation routines and the Laplace approximation.
Although this approach has favourable computational properties, we have found the inferences generated for the Naomi model to sometimes be inaccurate.
This has motivated us to look for a more accurate approach, which is flexible enough to be compatible with the model and fast enough to be run in production by country teams.

To obtain fast, accurate inferences for the Naomi model we develop a new inference methodology which combines the simplified integrated nested Laplace approximation (INLA) approach of @wood2020simplified with adaptive Gauss-Hermite quadrature (AGHQ).
INLA is an approach to approximate Bayesian inference based on nested Laplace approximations and numerical quadrature.
The key innovation of @rue2009approximate is an approximation which enables accurate latent field posterior marginals without explicitly computing the full Laplace approximation for each element.
Simplified INLA [@wood2020simplified] extends INLA by relaxing the sparsity assumptions on the latent field required for this approximation to be accurate.
This extension facilitates inference for models like Naomi, which fall within the extended latent Gaussian model (ELGM) [@stringer2022fast] class, and were not previously amenable to inference with INLA.
ELGMs build on latent Gaussian models (LGMs) by allowing each element of the linear predictor to depend on any subset of elements from the latent field.
We combine simplified INLA with AGHQ, a quadrature rule based on the theory of polynomial interpolation which adapts to the integrand based on the Hessian at the mode.
Though no theory yet exists for the nested case, the first stochastic convergence results for adaptive quadrature rules were recently obtained by @bilodeau2021stochastic using AGHQ.
We implement our method as an extension of the `aghq` \textsf{R} package [@stringer2021implementing].
Since `aghq` is designed to naturally interface with `TMB`, use of our method is simple when provided a \textsf{C++} user template for the log-posterior.

The remainder of this paper is organised as follows.
Section \ref{sec:naomi} describes a simplified version of the Naomi model that we consider in this paper.
Section \ref{sec:fastinferencemethods} outlines our approach to fast, accurate Bayesian inference using simplified INLA and AGHQ.
As a case-study, we compare the accuracy of our inference method to `TMB` and `tmbstan` for the simplified Naomi model fit to data from Malawi, in Section \ref{sec:naomi}.
We also demonstrate a Bayesian workflow, illustrating the applicability of these tools in a deterministic inference setting.
Finally, in Section \ref{sec:conclusions} we discuss our conclusions, how we anticipate our method might be useful for other models, and directions for future research.

@@

@eaton2021naomi specify a joint model linking small-area estimation models of HIV prevalence from household surveys, HIV prevalence among pregnant women from antenatal care clinics, and antiretroviral therapy (ART) coverage from routine health data collection.
The full model is defined over three time points: $T_1$ the time of the most recent household survey with HIV testing; $T_2$, the current time period; and $T_3$, a short term projection period.
We consider a simplified version of the model which is defined only at $T_1$ omitting nowcasting of $T_2$ and temporal projection to $T_3$.
Below we provide an overview of the three components in the simplified model, highlighting the aspects which make it a challenge for existing inferential approaches.
A more complete mathematical description of the simplified model, as well as a \textsf{C++} template for the log-posterior, are provided in the appendix.

@@

Consider a country in sub-Saharan Africa where a household survey has taken place during at time $T_1$.
Let $x$ index district, $a$ five-year age band, and $s$ sex and.
For ease of notation, let $i$ index the finest district-age-sex division included in the model.
The data we observe may be aggregated over indices $i$, so we let $\mathcal{I}$ be a set of $i$ for which observations are reported.
Let $N_i \in \mathbb{N}$ be the population size, $\rho_i \in [0, 1]$ be HIV prevalence, $\alpha_i \in [0, 1]$ be ART coverage, and $\lambda_i > 0$ be the annual HIV incidence rate.
We specify independent mixed effects models for HIV prevalence and ART coverage in the general population on the logit scale such that
\begin{align*}
\text{logit}(\rho_i) &= \eta^\rho_i, \\
\text{logit}(\alpha_i) &= \eta^\alpha_i,
\end{align*}
for certain choice of linear predictors $\eta^\rho_i$ and $\eta^\alpha_i$.
For the HIV incidence rate we use a mixed effects model on the log scale
$$
\log(\lambda_i) = \eta^\lambda_i(\{\rho_i, \alpha_i\}_{i \in \mathcal{I}}),
$$
where the linear predictor depends on $\{\rho_i, \alpha_i\}_{i \in \mathcal{I}}$ for some $\mathcal{I}$.
Let $\kappa_i$ be the proportion recently infected among HIV positive persons.
For each set of observed strata indices $\mathcal{I}$, we calculate the weighted observations $\hat \theta_\mathcal{I}$ for $\theta \in \{\rho, \alpha, \kappa\}$ with respective Kish effective sample sizes 
$$
M^{\hat \theta}_\mathcal{I} = \frac{\left(\sum_j w_j\right)^2}{\sum_j w_j^2},
$$
where $j$ indexes individuals across all strata $i \in \mathcal{I}$, with corresponding survey weights $w_j$.
The observed number of indicator cases is then
$$
Y^{\hat \theta}_{\mathcal{I}} = M^{\hat \theta}_{\mathcal{I}} \cdot \hat \theta_{\mathcal{I}}.
$$
For $\theta \in \{\rho, \alpha, \kappa\}$ we model these aggregate observations using a binomial working likelihood
$$
Y^{\hat \theta}_{\mathcal{I}} \sim \text{xBin}(M^{\hat \theta}_{\mathcal{I}}, \theta_{\mathcal{I}}),
$$
where $\theta_{\mathcal{I}}$ are the following appropriately weighted aggregates
\begin{align*}
\rho_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i}{\sum_{i \in \mathcal{I}} N_i}, \\
\alpha_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \alpha_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}, \\
\kappa_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} N_i \rho_i \kappa_i}{\sum_{i \in \mathcal{I}} N_i \rho_i}.
\end{align*}
To link the proportion recently infected among HIV positive persons $\kappa_i$ to HIV incidence $\lambda_i$ we use
$$
\kappa_i = 1- \exp \left( - \lambda \cdot \frac{1 - \rho_i}{\rho_i} \cdot (\Omega_T - \beta_T) - \beta_T \right),
$$
where the mean duration of recent infection $\Omega_T$ and false recent ratio $\beta_T$ are strongly informed by priors for the particular survey.

@@

We model HIV prevalence $\rho^\text{ANC}_i$ and ART coverage $\alpha^\text{ANC}_i$ among pregnant women as being offset on the logit scale from the general population indicator as follows
\begin{align*}
\text{logit}(\rho^\text{ANC}_i) &= \text{logit}(\rho_i) + \eta^{\rho^\text{ANC}}_i, \\
\text{logit}(\alpha^\text{ANC}_i) &= \text{logit}(\alpha_i) + \eta^{\rho^\text{ANC}}_i.
\end{align*}
We inform these processes by specifying likelihoods for the following aggregate ANC data from the year of the most recent survey: the number of ANC clients with ascertained status $X^\text{ANC}_\mathcal{I}$, the number of those with positive status $Y^\text{ANC}_\mathcal{I}$, and the number of ANC clients already on ART prior to their first ANC visit $Z^\text{ANC}_\mathcal{I}$.
We use the binomial working likelihoods
\begin{align*}
Y^\text{ANC}_\mathcal{I} &\sim \text{Bin}(Y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}) \\
Z^\text{ANC}_\mathcal{I} &\sim \text{Bin}(Y^\text{ANC}_\mathcal{I}, \alpha^\text{ANC}_{\mathcal{I}}),
\end{align*}
where again we use weighted aggregates
\begin{align*}
\rho^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}}{\sum_{i \in \mathcal{I}} \Psi_i}, \\
\alpha^\text{ANC}_{\mathcal{I}} = \frac{\sum_{i \in \mathcal{I}} \Psi_i \rho^\text{ANC} \alpha^\text{ANC}_i}{\sum_{i \in \mathcal{I}} \Psi_i \rho_i^\text{ANC}},
\end{align*}
with $\Psi_i$ the number of pregnant women.

@@

<!-- To-do: define what exactly $p(\y, \x, \btheta)$ is -->
Given a \textsf{C++} user template `model.cpp` for the negative unnormalised log posterior $- \log p(\y, \x, \btheta)$, we obtain the posterior marginal approximations $\{ \tilde p(x_i \, | \, \y) \}_{i = 1}^n$ and $\tilde p(\theta_j \, | \, \y)_{j = 1}^m$ via the following algorithm, comprised of nested applications of Laplace approximation and adaptive Gauss-Hermite quadrature.

@@

1. Calculate the mode, Hessian at the mode, and lower Cholesky
\begin{align*}
\hat{\btheta} &= \argmax_{\btheta} {\tilde p_\text{LA}(\btheta, \y)}, \\
\mathbf{H} &= \frac{\partial^2}{\partial \btheta \partial \btheta^\top} - \log \tilde p_\text{LA}(\btheta, \y) \rvert_{\btheta = \hat \btheta} \\
\mathbf{H}^{-1} &= \mathbf{L} \mathbf{L}^\top,
\end{align*}
of the Laplace approximation
$$
\tilde p_\text{LA}(\btheta, \y) = \frac{p(\y, \x, \btheta)}{\tilde p_\text{G}(\x \, | \, \btheta, \y)} \Big\rvert_{\x = \hat \x(\btheta)}
$$
where $\tilde p_\text{G}(\x \, | \, \btheta, \y) = \mathcal{N}(\x \, | \, \hat \x(\btheta), \mathbf{H}(\btheta)^{-1})$ is a Gaussian approximation to $p(\x \, | \, \btheta, \y)$ with mode and precision matrix given by
\begin{align*}
\hat \x(\btheta) &= \argmin_\x - \log p(\y, \x, \btheta), \\
\mathbf{H}(\btheta) &= \frac{\partial^2}{\partial x \partial x^\top} - \log p(\y, \x, \btheta) \rvert_{\x = \hat \x(\btheta)}.
\end{align*}
2. Generate a set of nodes $\z \in \mathcal{Q}(m, k)$ and weights $\omega: \z \in \mathcal{Q}(m, k) \to \mathbb{R}$ from a Gauss-Hermite quadrature rule with $m = \dim(\btheta)$, and $k$ nodes per dimension, which are then adapted based on the mode and lower Choleksy via $\btheta(\z) = \hat{\btheta} + \mathbf{L} \z$.
We typically set $k = 3$ such that there are $3^m$ nodes in total.
3. Use this quadrature rule to calculate $\tilde p_{\text{AQ}}(\y)$ as follows
$$
\tilde p_{\text{AQ}}(\y) = \sum_{\z \in \mathcal{Q}(m, k)} \tilde p_\text{LA}(\btheta(\z), \y) \omega(\z).
$$
4. For $x_i \in \{\ldots\}$ and $\btheta \in \{ \btheta(\z) \}_{\z \in \mathcal{Q}(m, k)}$ calculate the modes and Hessians
\begin{align*}
\hat \x_{-i}(x_i, \btheta) &= \argmin_{\x_{-i}} - \log p(\y, x_i, \x_{-i}, \btheta), \\
\mathbf{H}_{-i, -i}(x_i, \btheta) &= \frac{\partial^2}{\partial \x_{-i} \partial \x_{-i}^\top} - \log p(\y, x_i, \x_{-i}, \btheta) \rvert_{\x_{-i} = \hat \x_{-i}(x_i, \btheta)},
\end{align*}
where optimisation to obtain $\hat \x_{-i}(x_i, \btheta)$ may be initialised at $\hat \x(\btheta)_{-i}$.

@@

* We developed an approximate Bayesian inference algorithm to solve a challenging problem in the small-area estimation of HIV in low resource settings
* Following further testing, we anticipate adding our inference method as an option for 
* The flexibility of our method implementation, including compatibility with any `TMB` \textsc{C++} template, allows broader use, as well as investigation of, deterministic inference methods than had previously been possible. We are excited about exploration of its accuracy for the following challenging model structures which could not previously be fit: list of model types here.
* Our approach is similar to that of `inlabru` [@bachl2019inlabru] in the following ways, and different in the following other ways
* We demonstrated a Bayesian workflow for deterministic inference methods
